\begin{exercises}
		% Topics:
		% 
	\begin{problist}
	\prob Check that curves of the form $x^2 + y^2 = C$ satisfy the differential equation $\dfrac{dy}{dx} = -\dfrac{x}{y}$.
	
	
	\prob Is the piecewise-defined function
	$$
	y(x) = \begin{cases}
 		-x^2 & \text{ if } x< 0 \\
 		x^2 & \text{ if } x \geq 0		
	\end{cases}
	$$
	a solution of the differential equation $xy'-2y=0$ on $(-\infty,\infty)$?
	
	
	\prob Consider the differential equation
	$$ y'''' - 8y''' + 26 y'' - 40y'+25y=0.$$
	
	\begin{enumerate}
		\item Is $y=3e^{2x}$ a solution?
		\item Is $y=4 e^{2x}\sin(x)$ a solution?
		\item Is $y=-8 x e^{2x}\cos(x)$ a solution?
		\item For the three function above, if they are solutions, what are initial conditions of the form
			\begin{itemize}
				 \item[] $y(0) =$
				 \item[] $y'(0) =$
				 \item[] $y''(0) =$
				 \item[] $y'''(0) =$
			\end{itemize}
			that the solution satisfies?
	\end{enumerate}


	\prob Consider the functions
	\begin{align*}
		f(x) & = 3x + x^2 	& g(x) & = e^{-7x} \\
		h(x) & = \sin(x) 	& j(x) & = \sqrt{x} \\
		k(x) & = 8 e^{3x}	& \ell(x) & = -2 \cos(x)
	\end{align*}
	
	Match each differential to one or more functions which are solutions.
	
	\begin{enumerate}
		\item $y'=3y$
		\item $y''+9y'+14y=0$
		\item $y''+y=0$
		\item $2x^2y'' + 3xy'=y$
	\end{enumerate}
	
	
	
	\prob Consider the differential equation $u' = -2(u-10)$.
	
	\begin{enumerate}
		\item Check that the curves of the form $u = 10 + C e^{-2t}$ satisfy the differential equation.
		\item Sketch one solution of the differential equation.
		\item Sketch all the integral curves for the differential equation.
		\item What is the difference between a solution passing through the point $(1,20)$ and an integral curve passing through the same point?
	\end{enumerate}


	\prob Consider the differential equation $y'\big( 3y^2-1\big) = 1$.
	
	\begin{enumerate}
		\item Check that the curves of the form $y^3-y=x+C$ satisfy the differential equation.
		\item Sketch the solution of the differential equation that passes through $(1,1)$.
		\item Sketch the integral curve for the differential equation that passes through $(1,1)$.
		\item What is the difference between a solution passing through the point $(1,1)$ and an integral curve passing through the same point?
		\item Repeat (b)--(d) with the points $(1,0)$ and $(1,-1)$ instead of $(1,1)$.
	\end{enumerate}


	\prob Consider the ODE $y'(t) = \big(y(t)\big)^2$.
	One of these two graphs {\bf cannot} describe the solution. 
	Which one? 
	
	
	\begin{center}
	\begin{tikzpicture}
		\draw[-{\seta}] (-1,0) -- (3,0) node[above] {$t$};
		\draw[-{\seta}] (0,-3) -- (0,1) node[left] {$y$};
		\draw[ultra thick,domain=0.5:2.5,smooth,variable=\x,blue] plot ({\x},{(\x*\x-5)/3-1});
	\end{tikzpicture}
	\hfil
	\begin{tikzpicture}
		\draw[-{\seta}] (-1,0) -- (3,0) node[above] {$t$};
		\draw[-{\seta}] (0,-3) -- (0,1) node[left] {$y$};
		\draw[ultra thick,domain=0.5:2.5,smooth,variable=\x,blue] plot ({\x},{-((\x-3.5)^2)/4-0.25});
	\end{tikzpicture}	
	\end{center}

	\prob We seek a first-order ordinary differential equation \quad $y' = f(y)$ \quad whose solutions satisfy
	$$
	\begin{cases}
	y(x)  \mbox{ is concave up if } y < 1 \\
	y(x) \mbox{ is concave down if } y > 1
	\end{cases}
	$$
	%
	Write down or graph a function $f(y)$ that would produce such solutions.

	
	\end{problist}
\end{exercises}
