\newpage

\begin{exercises}

	\begin{problist}
	\prob Show that all autonomous differential equations are separable.
	
	\prob Consider the ODE $y'=\displaystyle \frac1y$.
	Then which of the statements below are true or false and justify your choice.
	\begin{enumerate}
		\item The solutions always stay positive.
		\item The solutions always stay negative.
		\item The solutions never change sign.
		\item The solutions always change sign.
		\item Some solutions change sign, and some don't.
	\end{enumerate}
	
	\prob In this problem, we are trying to deduce some properties of slope fields.
	\begin{enumerate}
		\item Sketch a slope field for the autonomous ODE $$y'=\cos(y).$$
		\item Sketch a slope field for the autonomous ODE $$y'=\tan(y).$$
		\item What is a common property of all slope fields of autonomous ODEs?
	\end{enumerate}	
	
	\prob Consider the ODE
	$$ 
	y' = {\rm sign}(y) = 
		\begin{cases}
			1 & \text{ if } y > 0 \\	
			0 & \text{ if } y = 0 \\	
			-1 & \text{ if } y < 0
		\end{cases}. 
	$$

	\begin{enumerate}
		\item What are the equilibrium solutions?
		\item Find two solutions that satisfy $y(0)=0$.
		\item Are there solutions that satisfy $\displaystyle \lim_{t \to \infty} y(t) = \pi$?
		\item What are the possible limits of solutions as $t\to \infty$?
	\end{enumerate}
	
%		y=0 sol
%		y=max(0,t+c) sol
%		y=min(0,-t+c) sol




	\prob \label{mod14-q-stable}Consider a function $f(y)$ such that
	\begin{itemize}
		\item $f(1)=0$;
		\item $f'$ is continuous for all $y$;
		\item $f'(1)<0$.
	\end{itemize}
	\begin{enumerate}
		\item Show that there is an open interval $(a,b)$ satisfying $1 \in (a,b)$ and $f'(y)<0$ for all $y \in (a,b)$.
		\item Show that %there is an interval $(c,d) \subseteq (a,b)$ such that
		\begin{itemize}
%			\item $1 \in (c,d)$, 
			\item $f(y)>0$ if $y \in (a,1)$, %(c,1)$,
			\item $f(y)<0$ if $y \in (1,b)$. %(1,d)$.
		\end{itemize}
		\item Consider the initial-value problem $y'=f(y)$ with $y(0)=y_0$. Show that this problem has a unique solution.
		\item Show that if $y_0 \in (a,b)$, then $\displaystyle \lim_{t \to \infty} y(t) = 1$.
		\item Write a Theorem about equilibrium points based on the results of this question.
	\end{enumerate}


	\prob Consider a function $f(y)$ such that
	\begin{itemize}
		\item $f(2)=0$;
		\item $f'$ is continuous for all $y$;
		\item $f'(2)>0$.
	\end{itemize}
	
	Complete a study similar to question \ref{mod14-q-stable} for this function $f$.
	
	
	
	\prob Consider an autonomous ODE $y'=f(y)$ with two stable equilibrium solutions $y=1$ and $y=2$ and where $f$ is continuous.
	\begin{enumerate}
		\item Show that there must exist another equilibrium point $c \in (1,2)$.
		\item Show that if $f(y)\not\equiv 0$ in $(1,2)$, then there must exist an equilibrium point $c \in (1,2)$ that is either semi-stable or unstable.
	\end{enumerate}

	\end{problist}
\end{exercises}