\begin{exercises}

	\begin{problist}
	\prob \label{mod17-gensol}Find the general solution of the problem $\vec{r}'(t) = A \vec{r}(t)$ for the following matrices:
	\begin{enumerate}
	\begin{minipage}{.2\textwidth}
		\item $A = \begin{bmatrix} -7 & 6 \\ -9 & 8 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 22 & 24 \\ -15 & -16\end{bmatrix}$;
		\item $A = \begin{bmatrix} 0 & 1 \\ -5 & 0 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 0 & 1 \\ 5 & 0 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 1 & \sqrt{3} \\ \sqrt{3} & -1\end{bmatrix}$;
		\item $A = \begin{bmatrix} 1 & \sqrt{3} \\ -\sqrt{3} & 1\end{bmatrix}$;
	\end{minipage}
	\qquad
	\begin{minipage}{.2\textwidth}
		\item $A = \begin{bmatrix} 0 & 1 \\ -4 & -4 \end{bmatrix}$;
		\item $A = \begin{bmatrix} -4 & -6 \\ 2 & 3 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 2 & -3 \\ 0 & 2 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 2 & 0 \\ 0 & 2 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 0 & 0 \\ 1 & 0\end{bmatrix}$; 
		\item $A = \begin{bmatrix} 0 & 0 \\ 0 & 1\end{bmatrix}$; 
	\end{minipage}
	\end{enumerate}

	\prob For each of the problems in the previous exercise, find the solution that satisfies the initial conditions:
	\begin{enumerate}[label=(\roman*)]
		\item $\vec{r}(0)=\begin{bmatrix} 0 \\ 0 \end{bmatrix}$;
		\item $\vec{r}(0)=\begin{bmatrix} 1 \\ 3 \end{bmatrix}$;
		\item $\vec{r}(1)=\begin{bmatrix} -2 \\ 2 \end{bmatrix}$.
	\end{enumerate}


	\prob Consider the problem $\vec{r}'(t) = \begin{bmatrix} 2 & 0 \\ 1 & 3 \end{bmatrix} \vec{r}(t) + \begin{bmatrix} -2 \\ 11 \end{bmatrix}$.
	\begin{enumerate}
		\item Show that $\vec{e}(t) = \begin{bmatrix} 1 \\ -4 	\end{bmatrix}$ is a solution of this problem.
		\item Find the general solution of 
		$$\vec{u}'(t) = \begin{bmatrix} 2 & 0 \\ 1 & 3 \end{bmatrix} \vec{u}(t).$$
		\item Let $\vec{r}(t) = \vec{u}(t) + \vec{e}(t)$. Show that this is a solution of the original problem.
		\item Let $\vec{r}_1(t)$ and $\vec{r}_2(t)$ be two solutions of the original problem. 
		\begin{enumerate}
			\item Is $\vec{r}_1(t) + \vec{r_2}(t)$ a solution? 
			\item Is $3\vec{r}_1(t) $ a solution? 
			\item Write a result on how one can safely combine solutions of non-homogeneous problems.
		\end{enumerate}
	\end{enumerate}

	
	\prob \label{prob:sys-nonhomogeneous}Consider the problem  \quad $\vec{r}'(t) = \begin{bmatrix} 1 & 2 \\ 3 & 0 \end{bmatrix}
 \vec{r}(t)+\begin{bmatrix} -5 \\ 3 \end{bmatrix}
$.
	\begin{enumerate}
		\item Observe that this system is an autonomous system of ODEs. What is the equilibrium solution? 
		\item Sketch its phase portrait using WolfamAlpha's streamplot function. What can you observe from the phase portrait?
		\item ``Redefine the centre'': Let the equilibrium solution solution you just found be called $\vec{e}$. Consider $\vec{u}(t) = \vec{r}(t) - \vec{e}$, where $\vec{r}(t)$ is the solution of the original problem. Show that 
			$$ \vec{u}'(t) = A \, \vec{u}(t).$$
		\item Find $\vec{u}(t)$.
		\item Find $\vec{r}(t)$.
		\item Write a procedure to solve any problem of the form
			$$ \vec{r}(t) = A \, \vec{r}(t) + \vec{b}. $$
	\end{enumerate}
	
	\prob Consider the problem $
\frac{d \,\vec{r}}{dt} = \begin{bmatrix} 4 & -1 \\ 8 & -2\end{bmatrix} \vec{r} - \begin{bmatrix} 5 \\ 10 \end{bmatrix}$.

	\begin{enumerate}
		\item Find $\vec{r}(t)$.
		\item What is the solution taht satisfies $\vec{r}(0) = \begin{bmatrix} 2\\3 \end{bmatrix}$?
		\item Sketch the phase portrait.
	\end{enumerate}

	
	\prob \label{prob:sys-superposition}Consider the problem \quad $\vec{r}'(t) = A \, \vec{r}(t)$.
	Assume that we have two solutions $\vec{r_1}(t)$ and $\vec{r_2}(t)$.
	\begin{enumerate}
		\item Show that $\vec{r}(t) = \vec{r_1}(t) + \vec{r_2}(t)$ is a solution also.
		\item Show that $\vec{r}(t) = 2\vec{r_1}(t) - 3\vec{r_2}(t)$ is a solution also.
		\item Find all possible solutions of the problem.
	\end{enumerate}
	
	
		\prob Consider the problem $\vec{r}'(t) = \begin{bmatrix} 1 & -2 \\ -2 & 1 \end{bmatrix} \vec{r}(t)$.
		\begin{enumerate}
			\item Find the solution that satisfies the initial condition $\vec{r}(0)=\begin{bmatrix}1 \\ 0\end{bmatrix}$. Call it $\vec{u}(t)$.
			\item Find the solution that satisfies the initial condition $\vec{r}(0)=\begin{bmatrix}0 \\ 1\end{bmatrix}$. Call it $\vec{v}(t)$.
			\item Define the matrix function
			$$ \Phi(t) = \begin{bmatrix} \vec{u}(t) \; | \; \vec{v}(t) \end{bmatrix} = \begin{bmatrix} u_1(t) & v_1(t) \\ u_2(t) & v_2(t) \end{bmatrix}.$$
			
			Show that $\vec{r}(t) = \Phi(t) \vec{r}_0$ is a solution of the original system of ODEs. Which initial condition does it satisfy?
			
			\item Write a result relating $\Phi(t)$ to the solution of initial-value problems.
			\end{enumerate}

	
	
		\prob \label{mod17:prob-W1}Consider a system of ODEs $\vec{r}'(t) = A \vec{r}(t)$ with two solutions $\vec{r}_1(t)$ and $\vec{r}_2(t)$. 
		
		We want to study the conditions that are necessary on the solutions $\vec{r}_1$ and $\vec{r}_2$ to guarantee that we can solve any initial-value problem.
		
		\begin{enumerate}
			\item What is the general solution for this problem?
			\item If the initial condition is $\vec{r}(0)= \begin{bmatrix} 1 \\ 2 \end{bmatrix}$, then what are the conditions on $\vec{r}_1,\vec{r}_2$ ?
			\item If the initial condition is $\vec{r}(0)= \vec{r_0}$, then what are the conditions on $\vec{r}_1,\vec{r}_2$ ?
		\end{enumerate}

		\prob Consider a system of ODEs $\vec{r}'(t) = A \vec{r}(t)$ with two solutions $\vec{r}_1(t), \vec{r}_2(t)$.
		
			Let $R(t)$ be the matrix 
				$R(t) = \begin{bmatrix} \vec{r}_1(t) & | & \vec{r}_2(t)	\end{bmatrix} $ 
				and let $W(t) = \det R(t)$.
		
		\begin{enumerate}
			\item Show that $W(t)$ is a solution of $W' = (a_{11} + a_{22}) W$.
			\item Solve the ODE above to obtain an expression for $W(t)$.
			\item Show that $W(t)$ is either identically zero, or it's never zero. 
			\item Use this result to simplify your answer to problem \ref{mod17:prob-W1}(c).
		\end{enumerate}
	
	\end{problist}
\end{exercises}
