\begin{exercises}

	\begin{problist}
	\prob Solve the differential equation $\displaystyle \ln(t) y' + \frac{1}{t}y = 3$.
	
	\prob Decide whether the following differential equations are separable, first-order linear, both, or neither. If they are of one type, solve it.
	\begin{enumerate}
		\item $\displaystyle (t^2+4) y'(t) = \frac{2t}{y^2}$
		\item $\displaystyle \frac{1}{t^2} y'(t) = 2$
		\item $\displaystyle \frac{dy}{dx} = \sqrt{y} (x+1)^2$
		\item $\displaystyle y'(t) = t + y$
		\item $\displaystyle y'(t) = t + y^2$
		\item $\displaystyle y'(t) = \frac{t}{y}$
		\item $\displaystyle y'(t) = -\frac{1}{t} y$
		\item $\displaystyle y'(t) = 1 - 4t - \frac5t y$
		\item $\displaystyle y'(t) = 5t - 2t y$
		\item $\displaystyle y'(t) = 2 + \cos^2(y)$
		\item $\displaystyle e^{-t} y'(t) - e^{-t} y = 3e^{2t}$
	\end{enumerate}
	
	\prob Decide whether the following statements are true or false. Give an explanation or a counterexample.
	
	\begin{enumerate}
		\item There are differential equations that are both separable and first-order linear.
		\item There are differential equations that are separable, but are not first-order linear.
		\item There are differential equations that are first-order linear, but not separable.
		\item There are first-order differential equations that are neither separable nor linear.
		\item All first-order linear differential equations have solutions defined in the whole real line.
	\end{enumerate}
	
	
	\prob Consider the differential equation
	$$ y' - \frac{y}{2(x+4)} = \frac{1}{2(x+4)} $$
	\begin{enumerate}
		\item Find the general solution.
		\item Find the solution with initial condition $y(0)=-5$.
		\item What is the domain of the previous solution?
		\item Find the solution with initial condition $y(-5)=-5$.
	\end{enumerate}


	\prob Even though the following differential equation is not linear, find its general solution:
		$$ 2 \ln(x) e^{2y} y'(x) + \frac{e^{2y}}{x} = 4x^3.$$

	\prob You are consulting for the police on Bernardo's murder.

		These are the facts about the murder:
		\begin{enumerate}
		\item The body was found at 9am;
		\item The body was found with the temperature of $25^o$C (average temperature is $37^o$C);
		\item The victim measured 185cm tall (average is 176cm) and weighed 75kg (average is 80kg);
		\item The body was found in his living room, which measured $25m^2$, and the thermostat was set to $22^o$C;
		\end{enumerate}
		
		There are five suspects that were with the victim the previous night:
		\begin{itemize}
		\item Pedro (height 179cm, weight 82kg) met with the victim at 6pm-8pm;
		\item Jason (height 176cm, weight 65kg) met with the victim at 8pm-10pm;
		\item Mihnea (height 183, 83 kg) met with the victim at 10pm-midnight;
		\item Paco (height 172cm, weight 79kg) met with the victim at midnight-2am;
		\item Ricardo (height 178cm, weight 76kg) met with the victim at 2am-4am.
		\end{itemize}
		
%		Recall Newton's Law of Cooling: ``The temperature change is proportional to the temperature difference''. The average proportionality constant for a human being is $k = \dfrac{\ln 5}{8}$.\\		
		
		Who killed Bernardo? (Your answer should stand in a court of law!)
	\end{problist}
\end{exercises}
