\begin{exercises}

	\begin{problist}
	\prob For each matrix from practice problem \ref{mod17-gensol} from Module 17, sketch its phase portrait and label them as asymptotically stable or unstable. The system of ODEs is $\vec{r}'(t) = A \vec{r}(t)$ for the following matrices:
	\begin{enumerate}
	\begin{minipage}{.2\textwidth}
		\item $A = \begin{bmatrix} -7 & 6 \\ -9 & 8 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 22 & 24 \\ -15 & -16\end{bmatrix}$;
		\item $A = \begin{bmatrix} 0 & 1 \\ -5 & 0 \end{bmatrix}$
		
		This is called a centre, which is stable, but not asymptotically stable. Can you tell why?
		\item $A = \begin{bmatrix} 0 & 1 \\ 5 & 0 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 1 & \sqrt{3} \\ \sqrt{3} & -1\end{bmatrix}$;
		\item $A = \begin{bmatrix} 1 & \sqrt{3} \\ -\sqrt{3} & 1\end{bmatrix}$;
	\end{minipage}
	\qquad
	\begin{minipage}{.2\textwidth}
		\item $A = \begin{bmatrix} 0 & 1 \\ -4 & -4 \end{bmatrix}$
		
		This is called an improper node.
		\item $A = \begin{bmatrix} -4 & -6 \\ 2 & 3 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 2 & -3 \\ 0 & 2 \end{bmatrix}$;
		\item $A = \begin{bmatrix} 2 & 0 \\ 0 & 2 \end{bmatrix}$;
		
		This is called a proper node.
		\item $A = \begin{bmatrix} 0 & 0 \\ 1 & 0\end{bmatrix}$; 
		\item $A = \begin{bmatrix} 0 & 0 \\ 0 & 1\end{bmatrix}$; 
	\end{minipage}
	\end{enumerate}
	
	\prob Consider a system of ODEs $\vec{r}'(t) = A \vec{r}(t)$.
	For each part, give an example of eigenvalues and eigenvectors of $A$ that would yield the required phase portrait:
	\begin{enumerate}
		\item Spiral sink (asymptotically stable);
		\item Spiral source (unstable);
		\item Centre (stable);
		\item Sink node (asymptotically stable);
		\item Source node (unstable);
		\item Saddle point (unstable);
		\item Improper node (stable);
		\item Improper node (unstable);
		\item Proper node (stable);
		\item Proper node (unstable);
	\end{enumerate}
	
	\prob Consider the system of ODEs
	$$
	\vec{r}'(t) = 
	\begin{bmatrix}
	1 & 1 \\ k & 1
	\end{bmatrix}\vec{r}(t).
	$$

	This system of ODEs changes behaviour depending on the parameter $k$.
	\begin{enumerate}
		\item Label the behaviour of the system for different values of $k$.
		\item We call the $k^\star$ the critical value of $k$ when the behaviour is different for $k<k^\star$ and for $k>k^\star$. For the critical value of $k$, sketch the phase portrait.
	\end{enumerate}
	
	
	\prob Consider the system of ODEs
	$$
	\vec{r}'(t) = 
	\begin{bmatrix}
	0 & 1 \\ -4 & -k
	\end{bmatrix}\vec{r}(t).
	$$

	This system of ODEs changes behaviour depending on the parameter $k$.
	\begin{enumerate}
		\item Label the behaviour of the system for different values of $k$.
		\item We call the $k^\star$ the critical value of $k$ when the behaviour is different for $k<k^\star$ and for $k>k^\star$. For the critical value of $k$, sketch the phase portrait.
		\item Which kinds of behaviours could be critical values?
	\end{enumerate}
	
	
	
	\prob Consider the system of ODEs
	$$
	\vec{r}'(t) = A \vec{r}(t).
	$$
	
	Let $T={\rm trace}(A) = a_{11}+a_{22}$, $D=\det(A) = a_{11}a_{22} - a_{12}a_{21}$, and $\Delta = T^2-4D$.
	
	\begin{enumerate}
		\item Show that the equilibrium solution is a saddle point if $D<0$.
		\item Show that the equilibrium solution is a spiral if $\Delta<0$ and $T\neq 0$.
		\item Show that the equilibrium solution is a centre if $T=0$ and $D>0$.
		\item When is the equilibrium point a node? \\

		\item Show that the equilibrium solution is asymptotically stable if $T<0$ and $D>0$.
		\item Show that the equilibrium solution is unstable if $T>0$ and $D<0$.
	\end{enumerate}






	\end{problist}
\end{exercises}
